Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-30 doc/article.tex
changes due to Dragan's comments

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib,verbatim}
\usepackage{grffile}
\usepackage{textcase}
\usepackage{hyperref}
\usepackage[overload]{textcase}

@@

\begin{document}

\thispagestyle{empty}
\pagenumbering{gobble} 
\begin{description}
\item[Title]Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys
\item[Thematic Science Area] Cosmology and Fundamental Physics
\item[Authors] and a list of authors
\item[Lead Author Contact Information]Alex Kim; Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720;
1-510-486-4621;
\href{mailto:agkim@lbl.gov}{agkim@lbl.gov}
\end{description}

\clearpage
\newpage
\setcounter{page}{0}
\pagenumbering{arabic}
\setcounter{page}{1}
\title{Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys}
\author[0000-0001-6315-8743]{A.~G.~Kim}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{G.~Aldering}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.~Antilogus}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{S.~BenZvi}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{H.~Courtois}
\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
Lyon, France} 
\author{T.~Davis}
\affiliation{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
\author{S.~Gontcho A Gontcho}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{R.~Graziani}
\affiliation{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
Physique de Clermont, F-63000 Clermont-Ferrand, France}    
\author{C.~Harper}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{C.~Howlett}
\affiliation{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
\author{D.~Huterer}
\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
Arbor, MI 48109, USA }
\author{C.~Ju}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.-F.~Leget}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{E.~V.~Linder}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.~McDonald}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{J.~Nordin}
\affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
\author{S.~Perlmutter}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\affiliation{Department of Physics, University of California Berkeley, 366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}
\author{N.~Regnault}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{M.~Rigault}
\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
Lyon, France} 
\author{A.~Slo\u{z}ar}
\affiliation{Brookhaven National Laboratory, Physics Department, Upton, NY
11973, USA} 
\author{others}

@@

In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
To be precise, the peculiar
velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
}
The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
follow the relation
$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
commonly used in place of $D$ to normalize the
overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.}

@@

Given these two advances, supernovae discovered by wide-field searches in the next decade will measure the growth of structure.
For example, 
over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
2.2\% uncertainty on fD  within this interval , where at
$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.

\section{Testing Gravity with Peculiar Velocity Surveys}
While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest here is in probing gravity
so we here focus  on $\gamma$.
To illustrate the distinction,
 $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega_M + \int \Omega_M^{\gamma} \ln{\Omega_M}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
 in $\Lambda$CDM;  two surveys with the same fractional precision  in $fD$ will have different precision in $\gamma$,
 with the one at lower redshift providing the tighter constraint. 
We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
The Fisher information matrix is
\begin{align}
F_{ij} 
%& = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
\label{fisher:eqn}
\end{align}
where
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{\delta \delta}(k,\mu) + \frac{1}{n} &
   P_{v\delta}(k,\mu)  \\
   P_{v\delta}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}
\label{cov:eq}
\end{equation}
and the two parameters considered are $\lambda \in \{\gamma,bD\}$.
The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
but are not considered.

@@

All the surveys considered  provide meaningful tests of gravity.
The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $>3\sigma$ level.
The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
Figure~\ref{var:fig} also shows uncertainties based on two other analyses, one that only uses peculiar velocities, and
one that uses both RSD and peculiar velocities independently.
Peculiar velocities alone
can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  $\sim 20$\%.
The implication is that there are important
$k$-modes that are  sample variance limited either in  overdensity and/or peculiar velocity who benefit from the
sample-noise suppression engendered by cross-correlations.

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/var.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
both RSD and peculiar velocities independently (dotted), and one  that uses
both RSD, peculiar velocities, and their cross-correlation (solid).
\label{var:fig}}
\end{figure}

Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
The top panel of Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{src/dvardz.png}\\
\includegraphics[width=0.4\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.4\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.4\textwidth]{src/dvardkmax.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Top: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ each as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] 
Increasing the survey duration  increases the number density of discovered supernovae. 
The increased survey duration improves the
precision in $\gamma$ as
shot-noise is relevant for all the surveys considered. The surveys with the highest fractional benefit are those that are more susceptible to shot noise:
those that monitor higher redshifts that have higher velocity dispersion, and shorter surveys  due to their lower number 
densities.
These trends are shown in the bottom-left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{max}$ for two- and ten-year surveys.
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has a similar relationship with $\sigma_\gamma$.
This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ shown in the bottom-right panel of Figure~\ref{dvar:fig}.
 \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.  Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$,
 which  is survey-independent.
 \end{description}

>>>>

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib,verbatim}
%\usepackage{grffile}
%\usepackage{textcase}
\usepackage{hyperref}
%\usepackage[overload]{textcase}

@@

\begin{document}
%\raggedright
\begin{raggedright}
\huge
Astro2020 Science White Paper \linebreak

Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys \linebreak
\normalsize

\noindent \textbf{Thematic Areas:} \hspace*{60pt} $\square$ Planetary Systems \hspace*{10pt} $\square$ Star and Planet Formation \hspace*{20pt}\linebreak
$\square$ Formation and Evolution of Compact Objects \hspace*{31pt} $\blacksquare$ Cosmology and Fundamental Physics \linebreak
  $\square$  Stars and Stellar Evolution \hspace*{1pt} $\square$ Resolved Stellar Populations and their Environments \hspace*{40pt} \linebreak
  $\square$    Galaxy Evolution   \hspace*{45pt} $\square$             Multi-Messenger Astronomy and Astrophysics \hspace*{65pt} \linebreak

\textbf{Principal Author:}

Name:	
 Alex Kim\linebreak						
Institution:  
Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720 \linebreak
Email:  \url{agkim@lbl.gov}
 \linebreak
Phone:  1-510-486-4621
 \linebreak

\textbf{Co-authors:} (names and institutions)
  \linebreak

\textbf{Abstract:}
In the upcoming decade cadenced wide-field imaging surveys 
will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
hundreds of thousands.  The increase in the number density  and solid-angle coverage 
of SNe~Ia, in parallel with improvements in the standardization of
their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
In the next decade the peculiar velocities of
SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
\end{raggedright}
\pagebreak
\newpage
\setcounter{page}{0}
\pagenumbering{arabic}
\setcounter{page}{1}
\title{Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys}
\author[0000-0001-6315-8743]{A.~G.~Kim}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{G.~Aldering}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.~Antilogus}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{S.~BenZvi}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{H.~Courtois}
\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
Lyon, France} 
\author{T.~Davis}
\affiliation{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
\author{S.~Gontcho A Gontcho}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{R.~Graziani}
\affiliation{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
Physique de Clermont, F-63000 Clermont-Ferrand, France}    
\author{C.~Harper}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{C.~Howlett}
\affiliation{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
\author{D.~Huterer}
\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
Arbor, MI 48109, USA }
\author{C.~Ju}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.-F.~Leget}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{E.~V.~Linder}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{P.~McDonald}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{J.~Nordin}
\affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
\author{S.~Perlmutter}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\affiliation{Department of Physics, University of California Berkeley, 366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}
\author{N.~Regnault}
\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
\author{M.~Rigault}
\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
Lyon, France} 
\author{A.~Slo\u{z}ar}
\affiliation{Brookhaven National Laboratory, Physics Department, Upton, NY
11973, USA} 
\author{others}

@@

In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
To be precise, the peculiar
velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
}
The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravity follow the relation
$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively \citep[see][for a review of
these gravity models within a cosmological context]{HUTERER201523}.  Adopting this parameterization to model gravity, peculiar velocity
surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
commonly used in place of $D$ to normalize the
overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.}

@@

Given these two advances, supernovae discovered by wide-field searches in the next decade will measure the growth of structure.
For example, 
over the course of a decade a SN survey relying on LSST discoveries 
plus spectroscopic redshifts can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
2.2\% uncertainty on fD  within this interval , where at
$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.

\section{Testing Gravity with Peculiar Velocity Surveys}
While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest here is in probing gravity
so we here focus  on $\gamma$.
To illustrate the distinction,
 $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega_M + \int \Omega_M^{\gamma} \ln{\Omega_M}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
 in $\Lambda$CDM;  two surveys with the same fractional precision  in $fD$ will have different precision in $\gamma$,
 with the one at lower redshift providing the tighter constraint. 
 In this section, we focus on how well peculiar velocity surveys in the upcoming decade can measure $\gamma$ and the sensitivity of the measurement
 on survey parameter choices.

We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
The Fisher information matrix is
\begin{align}
F_{ij} 
%& = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
\label{fisher:eqn}
\end{align}
where
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{\delta \delta}(k,\mu) + \frac{1}{n} &
   P_{v\delta}(k,\mu)  \\
   P_{v\delta}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}
\label{cov:eq}
\end{equation}
and the two parameters considered are $\lambda \in \{\gamma,bD\}$.
The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
but are not considered.

@@

All the surveys considered  provide meaningful tests of gravity.
The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $>3\sigma$ level.
The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
Figure~\ref{var:fig} also shows uncertainties based on two other analyses, one that only uses peculiar velocities, and
one that combines independent RSD and peculiar velocity results.
Peculiar velocities alone
can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  $\sim 20$\%.
The implication is that there are important
$k$-modes that are  sample variance limited either in  overdensity and/or peculiar velocity who benefit from the
sample-noise suppression engendered by cross-correlations.

\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{src/var.png}
\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
both RSD and peculiar velocities independently (dotted), and one  that uses
both RSD, peculiar velocities, and their cross-correlation (solid).
\label{var:fig}}
\end{figure}

Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.

{\bf Solid Angle $\Omega$:}
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.

{\bf Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
The leftpanel of Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{src/dvardz.png}
\includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
 $z_{max}$ for two- (dashed) and ten-year (solid) surveys.
\label{dvar:fig}}
\end{figure}

{\bf Redshift Depth $z_{max}$:}
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.

{\bf Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
An increased survey duration accumulates more supernovae, decreasing shot noise and  increasing the 
precision in $\gamma$ for all the surveys considered.
The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution,
shorter surveys and those with higher $z_{max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
These trends are shown in the left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{max}$ for two- and ten-year surveys.
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has a similar relationship with $\sigma_\gamma$.
This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ also shown in the left panel of Figure~\ref{dvar:fig}.

{\bf Minimum length scale, maximum wavenumber $k_{max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.  Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$,
 which  is survey-independent.

The condition of sample variance--shot noise equality of a $k$-mode is now examined.
This is important in tuning survey design because, as mentioned above, the range of surveys of interest have different relative contributions of sample variance and shot noise.
The RSD measurement is sample variance limited, as within a few years supernova discoveries from
the  rest-frame volumetric rate of $\phi=7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$ make shot noise unimportant.
For peculiar velocities,
the  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
can be compared using  Figure~\ref{power:fig}.  For these survey scenarios, sample variance and shot-noise equality 
occurs at a $k$  within the range that carries the most signal; both sample variance and shot-noise limited modes contribute to the constraints.
If $k_{max}$ were to decrease moderately some of the surveys could become sample variance limited, whereas increases in $k_{max}$ introduce additional
 shot-limited modes.

\section{Elements of a Peculiar Velocity Survey}
A successful peculiar velocity survey must accomplish several things, not just measure distances and host-galaxy redshifts.  These elements are as follows:

{\bf SNe Must Be Discovered:} Cadenced wide-field imaging surveys have proven to be efficient SN~Ia discovery machines.  ZTF and LSST
are examples, but not the sole representatives, of current and future surveys that will generate high surface-densities of SNe~Ia $n$ over a large solid angle $\Omega$. 
For example, 
LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
